Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning with Multiple Stimuli

### 2023-09-08 manuscript_draft.tex
all the updates

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Power to the Stimuli: Not the Effect},
  pdfauthor={Erin M. Buchanan1 \& Other Folks as Per Order on Doc2},
  pdflang={en-EN},
  pdfkeywords={power, sampling, accuracy in parameter estimation},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{power, sampling, accuracy in parameter estimation}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Power to the Stimuli: Not the Effect}
\author{Erin M. Buchanan\textsuperscript{1} \& Other Folks as Per Order on Doc\textsuperscript{2}}
\date{}

\shorttitle{Title}

@@

The authors would like to thank K.D. Valentine for her assistance in formulation of correction scores.

@@

\abstract{%
We will add an abstract that explains that we are going to 1) talk about this cool procedure, 2) show some simulations that the procedure works, and 3) give two examples of the procedure in action. You should try it out!
}

@@

This part we will put in a real introduction. Things we need to discuss:

\begin{itemize}
\tightlist
\item
  Repeated measures data with items
\item
  Accuracy in Parameter Estimation
\item
  Adaptive sampling
\item
  MLM designs account for item heterogeneity but a lot of people anova
\end{itemize}

\hypertarget{simulating-sample-size}{%
\subsection{Simulating Sample Size}\label{simulating-sample-size}}

Using ideas from accuracy in parameter estimation, we suggest the following procedure to determine a sample size for each item:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  Use pilot data that closely resembles your intended data collection, on the same or similar items that will be used in the study. In this procedure, we will assume that the pilot data is representative of a larger population of sampled items that you intend to assess.
\item
  Calculate the standard error of each item from the pilot data to create a cutoff score for when an item is ``accurately measured''. The simulations below will explore what criterion to use when determining the cutoff score from the pilot data.
\item
  Sample, with replacement, from your pilot data using sample sizes starting at 20 participants and increase in small units (e.g., 20, 25, 30) up to a value that you consider the maximum sample size. We will demonstrate example maximum sample sizes based on the data simulation below; however, a practical maximum sample size may be determined by time (e.g, one semester data collection) or researcher resources (e.g., 200 participants worth of funding). While 20 participants would likely represent an underpowered study, we simply suggest this starting minimum for simulation purposes.
\item
  For each simulated sample, calculate the standard error for each item, and use these values to ascertain the percentage of items that meet the cutoff score determined in step 2.
\item
  Find the minimum sample size that meets 80\%, 85\%, 90\%, and 95\% of the items. We recommend these scores to ensure that most items are accurately measured, in a similar vein to common power criteria suggestions. Each researcher can determine which of these is their minimum or maximum sample size (e.g., individual can choose to use 80\% as a minimum and 90\% as a maximum).
\item
  Report these values, and designate a minimum sample size, the cutoff criterion, and the maximum sample size. Each researcher should also report if they plan to use an adaptive design, which would stop data collection after meeting the cutoff criterion for each item.
\end{enumerate}

\hypertarget{key-issues}{%
\subsection{Key Issues}\label{key-issues}}

Given the long history of research on power, there are a few key issues that this procedure should address:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  We should see differences in projected sample sizes based on the variability in the variance for those items (i.e., heterogeneity should increase projected sample size).
\item
  We should see projected sample sizes that ``level off'' when pilot data increases. As with regular power estimates, studies can be ``overpowered'' to detect an effect, and this same idea should be present. For example, if one has a 500 person pilot study, our simulations should suggest a point at which items are likely measured well, which may have happened well before 500.
\end{enumerate}

\hypertarget{method}{%
\section{Method}\label{method}}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man]{apa7}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\makeatletter
\usepackage{etoolbox}
\patchcmd{\maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{accuracy in parameter estimation, power, sampling, simulation, hypothesis testing}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning with Multiple Stimuli},
  pdfauthor={Erin M. Buchanan1 \& Other Folks as Per Order on Doc2},
  pdflang={en-EN},
  pdfkeywords={accuracy in parameter estimation, power, sampling, simulation, hypothesis testing},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning with Multiple Stimuli}
\author{Erin M. Buchanan\textsuperscript{1} \& Other Folks as Per Order on Doc\textsuperscript{2}}
\date{}

\shorttitle{AIPE BOOTSTRAPPING}

@@

@@

\abstract{%
The planning of the sample size for research studies often focuses on obtaining a significant result given a specified level of power, significance, and an anticipated effect size. This planning requires prior knowledge of the study design and a statistical analysis to calculate the proposed sample size. However, there may not be one specific testable analysis from which to derive power (Silberzahn et al., 2018) or a hypothesis to test for the project (e.g., creation of a stimuli database). Modern power and sample size planning suggestions include accuracy in parameter estimation (AIPE, Kelley, 2007; Maxwell et al., 2008) and simulation of proposed analyses (Chalmers \& Adkins, 2020). These toolkits provide flexibility in traditional power analyses that focus on the if-this, then-that approach, yet, both AIPE and simulation require either a specific parameter (e.g., mean, effect size, etc.) or statistical test for planning sample size. In this tutorial, we explore how AIPE and simulation approaches can be combined to accommodate studies that may not have a specific hypothesis test or wish to account for the potential of a multiverse of analyses. Specifically, the examples focus on studies that adopt multiple items and suggest that sample sizes can be planned to measure those items adequately and precisely, regardless of statistical test. We demonstrate that pilot data can be used to determine a sample size that represents well-measured data. This tutorial also provides multiple code vignettes that researchers can adapt and apply to their own measures.
}

@@

Statistical power and power analyses are arguably one of the most important components in planning a research study (Cohen, 1990). However, if reviews of transparency and openness in research publications are any clue, the social sciences have failed to fully implement power analyses as part of their common efforts (Hardwicke et al., 2020, 2022). The replication ``crisis'' and credibility revolution have shown that published studies in psychology are underpowered (Open Science Collaboration, 2015; Vazire, 2018). Pre-registration of a study involves outlining the study and hypotheses before data collection begins (Chambers et al., 2014; Nosek \& Lakens, 2014; Stewart et al., 2020), and details of a power analyses or other limitations on resources are often used to provide justification for the pre-registered sample quota. Given the combined issues of publish-or-perish and that most non-significant results do not result in published manuscripts, one may expect that power analysis would be especially critical for early career researchers (Rosenthal, 1979; Simmons et al., 2011). At best, an underpowered study provides limited insight (Halpern, 2002), and it can be difficult to know if a poorly implemented power analysis is better than no power analysis.

A recent review of power analyses found - across psychology journal articles - researchers did not provide enough information to understand their power analyses and often chose effect sizes that were unjustified (Beribisky, 2019). One solution to this power analysis problem is the plethora of tools made available for researchers to make power computations accessible to non-statisticians with the caveat that a solid education in power is necessary to use these tools. G*Power is one of the most popular free power software options (Erdfelder et al., 1996; Faul et al., 2007) that provides a simple point and click graphical user interface for power calculations (however, see Brysbaert, 2019). Web-based tools have also sprung up for overall and statistical test specific sample size planning including \url{https://powerandsamplesize.com} and \url{https://designingexperiments.com} (Anderson et al., 2017). \emph{R}-coding based packages, such as \emph{pwr} (Champely et al., 2017), \emph{faux} (DeBruine, 2021), and \emph{SimDesign} (Chalmers \& Adkins, 2020), can be used to examine power and sample size planning usually with simulation. Researchers must be careful using any toolkit, as errors can occur with the over-reliance on software (e.g., it should not be a substitute for critical thinking, Nuijten et al., 2016). When computing sample size estimates, it is important to remember that these values are estimations, not exact calculations guaranteed to produce a specific result (Batterham \& Atkinson, 2005). For example, if any parameter estimated by the researcher was not found in the study (i.e., a smaller effect size than used for the power analysis), then the sample size estimate tied to that specific level of power may be incorrect.

Changes in publication practices and research design have also created a new wrinkle in providing a sample size plan for a research study. While statistics courses often suggest that a specific research design leads to a specific statistical test, meta-science work has shown that given the same data and hypothesis, researchers can come up with multiple ways to analyze the data Coretta et al. (2023). Therefore, a single power analysis only supports the specific analysis in which the researcher expects to test. Analyses may evolve during the research project or be subject to secondary analysis; thus, power and sample size estimation based on one analysis is potentially less useful than previously imagined. Further, research projects often have multiple testable hypotheses, but it is unclear which hypothesis or test should be used to estimate sample size with a power analysis. Last, research publications may not even have a specific, testable hypothesis, as some publications are intended to curate a large dataset for future reuse (i.e., stimuli database creation, Buchanan et al., 2019).

In light of these analytical (or lack thereof) concerns, we propose a new method that combines accuracy in parameter estimation Maxwell et al. (2008) and bootstrapped simulation on pilot data (Rousselet et al., 2022). This method accounts for a potential lack of hypothesis test (or simply not a good way to estimate an effect size of interest), and/or an exploratory design with an unknown set of potential hypotheses and analytical choices. Additionally, we consider the nature of cognitive research designs that use multiple items to measure the phenomena of interest. For example, semantic priming is measured with multiple paired stimuli (Meyer \& Schvaneveldt, 1971), which traditionally was analyzed by creating person or item-level averages for an ANOVA (Brysbaert \& Stevens, 2018). However, the use of multilevel models with random effects for the stimuli used in a study have shown that we should be careful to assume that all items of a research study have the same ``effect'', as there is often variability in their impact on the outcomes of the study.

\hypertarget{accuracy-in-parameter-estimation}{%
\subsection{Accuracy in Parameter Estimation}\label{accuracy-in-parameter-estimation}}

AIPE shifts the focus away from finding a significant p-value to finding a parameter that is ``accurately measured''. For example, researchers may wish to detect a specific correlation in a study, \emph{r} = .35. They could then use AIPE to estimate the sample size needed to find a ``sufficiently narrow'' confidence interval around that correlation. Sufficiently narrow is often defined by the researcher using a minimum parameter size of interest and confidence intervals. Therefore, they could decide that their 95\% confidence interval should be approximately between .20 and .50, and sufficiently narrow could be defined as a width of .30 or .15 on each side. While confidence intervals are related to null hypothesis significance testing (i.e., 95\% confidence intervals that do not include zero would indicate a significant difference from zero at \(\alpha\) \textless{} .05), AIPE procedures suggest how we can define a sample size with a given width of confidence interval, regardless of whether it includes zero.

\hypertarget{bootstrapping-and-simulation}{%
\subsection{Bootstrapping and Simulation}\label{bootstrapping-and-simulation}}

Bootstrapping involves using data obtained to simulate similar datasets by drawing from the original data with replacement (Efron, 2000; Rousselet et al., 2022). Bootstrapping is a form of data simulation that allows one to calculate parameter estimates, confidence intervals, and more to simulate the potential population distribution, shape, and bias. Simulation is often paired with making up data for testing analyses and hypotheses based on proposed effect size or suggested population means. Generally, we would suggest starting with pilot data of a smaller sample size to understand the variability in potential items used to represent your phenomenon, especially if they are to be used in a larger study. However, given some background knowledge about the potential items, one could simulate example pilot data to use in. similar manner in our suggested procedure. Pilot or simulated data would be used to estimate the variability within items and select a ``sufficiently narrow'' window for item's confidence interval for AIPE (i.e., by selecting a specific standard error criterion, given the formula for confidence intervals). Bootstrapping would then be used to determine how many participants may be necessary to achieve a dataset wherein many items meet the required confidence interval.

\hypertarget{sequential-testing}{%
\subsection{Sequential Testing}\label{sequential-testing}}

Researchers could then use sequential testing to estimate their parameter of interest after each participant's data was collected to determine whether they have achieved their expected width of the confidence interval around that parameter. One would set a minimum sample size (i.e., based on known data collection ability) and use the confidence interval width as a stopping rule (i.e., stop data collection when the CI is narrow, as defined above). Next, researchers would use the estimated sample size associated with the simulation results of many items obtaining the stopping rule as a maximum sample size (i.e., we expect 90\% of items to meet our stopping rule with 100 participants based on simulation). By defining each of these components, researchers could ensure a feasible minimum sample size, a way to stop data collection when goals have been met, and a maximum sample size rule to ensure an actual end to data collection. Therefore, we proposed that we should be able to leverage the ideas behind AIPE, paired with simulation and bootstrapping, to estimate the minimum and maximum proposed sample sizes and stopping rules for repeated measures studies with expected variability in parameter estimates for items.

\hypertarget{proposed-method-to-calculate-sample-size}{%
\subsection{Proposed Method to Calculate Sample Size}\label{proposed-method-to-calculate-sample-size}}

Using these ideas, we suggest the following procedure to determine a sample size for each item:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  Use pilot data that closely resembles data you intend to collect. This dataset should contain items that are identical or similar to those that will be collected in the study. In this procedure, it is important to ensure that the data is representative of a larger population of sampled items that you intend to assess. Generally, pilot data sample sizes will be small, as the goal would be to determine how many items would be necessary to reach a ``stable'' standard error.
\item
  Calculate the standard error (SE) in the subset of pilot data associated with each item to create a cutoff score that defines when items are considered ``accurately measured''. Next, the simulations below will explore what criterion should be used to determine the cutoff score from the pilot data.
\item
  Sample, with replacement, from your pilot data using sample sizes starting at 20 participants and increase in small units up to a value that you consider the maximum sample size. We will demonstrate example maximum sample sizes based on the data simulation below; however, a practical maximum sample size may be determined by time (e.g., one semester data collection) or resources (e.g., 200 participants worth of funding). Although 20 participants would likely yield imprecise estimates, we suggest this starting minimum for simulation purposes.
\item
  For each simulation, calculate the standard error for each item, and use these values to determine the percentage of items that meet the cutoff score determined in Step 2.
\item
  Find the minimum sample size so that 80\%, 85\%, 90\%, and 95\% of the items meet the cutoff score and can be considered accurately measured. We recommend these scores to ensure that most items are accurately measured, in a similar vein to the common power-criterion suggestions. Each researcher can determine which of these is their minimum or maximum sample size (e.g., individuals can choose to use 80\% as a minimum and 90\% as a maximum or use values from Step 3 based on resources).
\item
  Report these values, and designate a minimum sample size, the cutoff criterion, and the maximum sample size. Each researcher should also report if they plan to use an adaptive design, which would stop data collection after meeting the cutoff criterion for each item.
\end{enumerate}

These steps are summarized in Table \ref{tab:table-summary}. We will first demonstrate the ideas behind the steps using open data (Balota et al., 2007; Brysbaert et al., 2014). This example will reveal a few areas of needed exploration for the steps. Next, we portray simulations for the proposed procedure and find solutions to streamline and improve the sample size estimation procedure. Finally, we include additional resources for researchers to use to implement the estimation procedure.

\caption{\label{tab:table-summary}Proposed Procedure for Powering Studies with Multiple Items}

\footnotesize{

\begin{tabular}{m{1cm}m{6cm}m{6cm}}
\toprule
Step & Proposed Steps & Updated Steps\\
\midrule
1 & Use representative pilot data. & Use representative pilot data.\\
2 & Calculate standard error of each of the items in the pilot data. Determine the appropriate SE for the stopping rule. & Calculate standard error of each of the items in the pilot data. Using the 40\%, determine the cutoff and stopping rule for the standard error of the items.\\
3 & Create bootstrapped samples of your pilot data starting with at least 20 participants up to a maximum number of participants. & Create bootstrapped samples of your pilot data starting with at least 20 participants up to a maximum number of participants.\\
4 & Calculate the standard error of each of the items in the bootstrapped data. From these scores, calculate the percent of items below the cutoff score from Step 2. & Calculate the standard error of each of the items in the bootstrapped data. From these scores, calculate the percent of items below the cutoff score from Step 2.\\
5 & Determine the sample size at which 80\%, 85\%, 90\%, 95\% of items are below the cutoff score. & Determine the sample size at which 80\%, 85\%, 90\%, 95\% of items are below the cutoff score. Use the correction formula to adjust your proposed sample size based on pilot data size, power, and percent variability.\\
6 & Report all values. Designate one as the minimum sample size, the cutoff score as the stopping rule for adaptive designs, and the maximum sample size. & Report all values. Designate one as the minimum sample size, the cutoff score as the stopping rule for adaptive designs, and the maximum sample size.\\
\bottomrule
\end{tabular}

\hypertarget{example}{%
\section{Example}\label{example}}

In this section, we provide two examples of the suggested procedure. The first example includes concreteness ratings from Brysbaert et al. (2014). Instructions given to participants denoted the difference between concrete (i.e., ``refers to something that exists in reality'') and abstract (i.e., ``something you cannot experience directly through your senses or actions'') terms. Participants were then asked to rate concreteness of terms using a 1 (\emph{abstract}) to 5 (\emph{concrete}) scale. This data represents a small scale dataset that could be used as pilot data for a study using concrete word ratings. The data is available at \url{https://osf.io/qpmf4/}.

The second dataset includes a large scale dataset with response latencies, the English Lexicon Project (Balota et al., 2007). The English Lexicon Project consists of lexical decision response latencies for English words. In a lexical decision task, participants simply select ``word'' for real words (e.g., \emph{dog}) and ``nonword'' for pseudowords (e.g., \emph{wug}). The trial level data is available here: \url{https://elexicon.wustl.edu/}. Critically, in each of these examples, the individual trial level data for each item is available to simulate and calculate standard errors on. Data that has been summarized could potentially be used, as long as the original standard deviations for each item were present. From the mean and standard deviation for each item, a simulated pilot dataset could be generated for estimating new sample sizes. All code to estimate sample sizes is provided on our OSF page, and this manuscript was created with a \emph{papaja} (Aust et al., 2022) formatted Rmarkdown document.

For this example, our researcher wants to determine the differences in response latencies for abstract and concrete words. They will select \emph{n} = 40 words from the rating data that are split evenly into abstract and concrete ends of the rating scale. In the experiment, each participant will rate the words for their concreteness, and then complete a lexical decision task with these words as the object of interest. Using both datasets, we can determine the sample size necessary to ensure adequately measured ratings and response latencies.

\emph{Step 1}. The concreteness ratings data includes 27250 concepts that were rated for their concreteness. We randomly selected \emph{n} = 20 abstract words (\(M_{Rating}\) \textless= 2) and \emph{n} = 20 concrete words (\(M_{Rating}\) \textgreater= 4). In the original study, not every participant rated every word, which created uneven sample sizes for each word. Further, participants were allowed to indicate they did not know a word, and this data was set to missing data. In our sample of 40 words, the average pilot sample size was 28.23 (\emph{SD} = 1.35), and we will use 28 as our pilot sample size for this example.

The ELP response latency data includes 27250 word-forms, 219 that are listed as non-words, and 27031 real words. We selected the same words as the concreteness subset selected above. The average pilot sample size for this random sample was 32.80 (\emph{SD} = 0.56), and \emph{n} = 33 will be our pilot size for this example.

\emph{Step 2}. Table \ref{tab:table-example} demonstrates the cutoff scores for deciles of the standard errors for the items for the concreteness ratings and lexical decision response latencies. A researcher could potentially pick any of these cutoffs or other percentage options not shown here. We will use simulation to determine the suggestion that best captures the balance of adequately powering our sample and feasibility.

\emph{Step 3-5}. The pilot data was then simulated, with replacement, with samples from 20 to 300 increasing in units of 5, separately for concreteness and lexical decision times (Step 3). The standard error of each item was calculated for the bootstrapped samples (Step 4), and the percentage of items below each potential cutoff was gathered (Step 5). The smallest sample size with at least 80\%, 85\%, 90\%, and 95\% of items below the cutoff are reported in Table \ref{tab:table-example} (Step 5).

\caption{\label{tab:table-example}Sample Size Estimates by Decile for Concreteness Example}
